% Run provenance for the single pre-registered study (campaigns C1/C2/C3).
% Re-keyed to the H1--H5 hypothesis set and the deposited campaign datasets.

\chapter{Run provenance}\label{app:provenance}

Every confirmatory number in Chapter~\ref{ch:results} traces to one of three
pre-registered campaigns, each deposited as a standalone dataset under its own
DOI \emph{before} its analysis was written
(\texttt{01-PREREGISTRATION.md}, \S Versioning). The campaigns share an
environment fingerprint, recorded in every session's \texttt{runMetadata} and in
each archive's \texttt{artifact-manifest.json} (written by
\texttt{archive\_run.py --strict}): Kubernetes \textbf{v1.28.6}, kube-proxy mode
\textbf{ipvs}, containerd 1.7.11, Calico CNI, eight 2-vCPU/4-GiB workers,
namespace \texttt{online-boutique}, git \texttt{dirty: false},
\texttt{provenanceWarnings: []}. The frozen pre-registration is deposited at DOI
\href{https://doi.org/10.5281/zenodo.20690836}{10.5281/zenodo.20690836} (pinned
to its freeze commit \texttt{20097c1}); all three campaign datasets
\texttt{continue} the concept DOI
\href{https://doi.org/10.5281/zenodo.20639145}{10.5281/zenodo.20639145} and
\texttt{supplement} the pre-registration.

\section{Campaigns}\label{sec:campaigns}

\begin{center}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{0.15\textwidth}p{0.08\textwidth}p{0.30\textwidth}p{0.06\textwidth}p{0.09\textwidth}p{0.13\textwidth}}
\toprule
\textbf{campaign} & \textbf{tests} & \textbf{design} & \textbf{sess.} & \textbf{data commit} & \textbf{DOI} \\
\midrule
C1 dose-response & H1, H4, H5 & complete-block $f \in \{0, 0.25, 0.5, 0.75, 1.0\}$, $r{=}1$, \texttt{pod-delete} churn; each session visits all five levels in seeded-random order, $5\times3 = 15$ iterations & 8 (s01--s08) & \texttt{2ec934f} & \href{https://doi.org/10.5281/zenodo.20690737}{20690737} \\
\addlinespace
C2 replication rescue & H3 & between-subjects $r \in \{1,3\} \times \text{mode} \in \{\text{packed}, \text{anti-affine}\}$, three non-degenerate cells (r1-packed, r3-packed, r3-anti-affine) $\times\,8$, \texttt{node-drain} at $f{=}0.5$ & 24 & \texttt{e533d5b} & \href{https://doi.org/10.5281/zenodo.20726729}{20726729} \\
\addlinespace
C3 placement + DNS & H2 & within-subjects $f \in \{0,1\}$ at $r{=}1$, crossed with \texttt{dnsCache} $\in \{\text{off}, \text{on}\}$ (NodeLocal DNSCache via pod \texttt{dnsConfig}); 7 matched cache-off/cache-on pairs, \texttt{pod-delete} churn, 3 iterations per placement & 14 & \texttt{2409f35} & \href{https://doi.org/10.5281/zenodo.20748970}{20748970} \\
\bottomrule
\end{tabular}
\end{center}

All three DOIs resolve to Zenodo dataset depositions (CC-BY-4.0, open), linked by
related identifiers so each \texttt{continues} the concept record,
\texttt{isSupplementTo} the frozen pre-registration, and \texttt{references} the
earlier campaign(s). C1 was collected on data commit \texttt{2ec934f}; its
analysis runs on \texttt{main} at or after \texttt{7d6943e}, so the data commit
predates the analysis tooling by design (deviation D-2026-06-14-01). C2 and C3
were collected and analyzed on a single strict-clean commit each.

\section{Claims \texorpdfstring{$\rightarrow$}{to} campaigns}\label{sec:claims-runs}

\begin{center}
\footnotesize
\begin{tabular}{p{0.42\textwidth}p{0.5\textwidth}}
\toprule
\textbf{Claim (Chapter~\ref{ch:results})} & \textbf{Campaign / analysis} \\
\midrule
\textbf{H1} dose-response, $+13.35\,\%$ $<$ $15\,\%$ SESOI (\S\ref{sec:h1}) & C1 (8 sessions); Page's $L$ trend test on per-session level medians \\
\addlinespace
\textbf{H2} placement-dependence (reversed) + DNS arm (\S\ref{sec:h2}) & C3 (7 cache-off + 7 cache-on sessions); paired drop contrast + within-spread cache comparison \\
\addlinespace
\textbf{H3} replication rescue under node-drain (\S\ref{sec:h3}) & C2 (24 sessions); ART-ANOVA $r\times\text{mode}$ interaction on trough depth and user-route error \\
\addlinespace
\textbf{H4} descriptive placement frontier (\S\ref{sec:h4}) & C1 (the five $r{=}1$ \texttt{pod-delete} placements; analysis-only, no additional runs) \\
\addlinespace
\textbf{H5} layered scorecard ICC, mechanism 0.994 vs naive aggregate 0.066 (\S\ref{sec:h5}) & C1 (condition-level test-retest reliability across the 8 sessions) \\
\addlinespace
\textbf{Family Holm capstone} (\S\ref{sec:family-holm}) & H1, H2, H3, H5 corrected together; each member's family-input $p$ read verbatim from its own driver \\
\addlinespace
\textbf{H6} direction transfer (\S\ref{sec:h6}) & not attempted --- the cluster ran the ipvs proxier, not the iptables mode the hypothesis presumes; de-scoped before collection, no runs \\
\bottomrule
\end{tabular}
\end{center}

\section{Integrity}\label{sec:integrity}

Each session directory carries \texttt{summary.json}, its per-condition raw JSON
(\texttt{f-*.json}), \texttt{charts/}, and an \texttt{artifact-manifest.json}
pinning run identity, git provenance, the SHA-256 of the exact scenario YAML
injected, and a SHA-256 of every file in the session. Every campaign tarball
additionally carries a root \texttt{SHA256SUMS}; C2's and C3's raw checksums are
pinned in the committed manifests \texttt{c2-roundrobin-manifest.sha256} and
\texttt{c3-dns-manifest.sha256} (computed pre-analysis). All sessions passed
\texttt{archive\_run.py --strict} (clean tree, scenario hashes present,
\texttt{runMetadata} present): C1 8/8, C2 24/24 (all accepted, untainted, 8 per
cell), C3 14/14 (all accepted, untainted, 7 per cache mode). Per-session
\texttt{summary.json} SHA-256 anchors are tabulated in each campaign's deposit
manifest; representative anchors are
C1~s01 \texttt{b2282cce\ldots f2301456},
C2~\#1 (r1-packed) \texttt{e4b3c50f\ldots 504f408}, and
C3~\#1 (cache-off) \texttt{83169d79\ldots f7fc6a82}.

\section{External-validity campaign (exploratory)}\label{sec:hotel-provenance}

The second-workload replication (\S\ref{sec:external-validity}) is
\textbf{exploratory and outside the frozen Holm family}; it is deposited and
provenance-gated to the same bar as the confirmatory campaigns but tests no
pre-registered hypothesis, so it is recorded separately here. Workload:
DeathStarBench \texttt{hotelReservation} (namespace \texttt{hotel-reservation}),
a deep frontend fan-out with per-service datastore pairs and Go gRPC over Consul
service discovery, on the same eight-worker cluster.

\begin{center}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{0.16\textwidth}p{0.08\textwidth}p{0.34\textwidth}p{0.06\textwidth}p{0.09\textwidth}p{0.11\textwidth}}
\toprule
\textbf{arm} & \textbf{test} & \textbf{design} & \textbf{sess.} & \textbf{data commit} & \textbf{DOI} \\
\midrule
C1-hotel dose-response & H1 & complete-block $f \in \{0,0.25,0.5,0.75,1.0\}$, $r{=}1$, \texttt{pod-delete} on the search path, $5\times3$ iterations & 8 & \texttt{bdf1ccb} & \href{https://doi.org/10.5281/zenodo.20792129}{20792129} \\
\addlinespace
C2-hotel replication rescue & H3 & three cells (r1-packed, r3-packed, r3-anti-affine) $\times\,8$, \texttt{node-drain} at $f{=}0.5$, round-robin assignment & 24 & \texttt{bdf1ccb} & (same) \\
\bottomrule
\end{tabular}
\end{center}

Both arms were collected and analyzed on \texttt{main} commit \texttt{bdf1ccb}
(\texttt{git.dirty: false}). All 32 sessions are \texttt{doctor --strict} clean;
1 of the 144 churn iterations (one iteration of C1-hotel order-seed 8, collected
after a deliberate pause/restart) carried an \texttt{app\_ready\_timeout} taint
and is excluded by the registered healthy-only rule. The raw data are deposited
as a single CC-BY-4.0 dataset, DOI
\href{https://doi.org/10.5281/zenodo.20792129}{10.5281/zenodo.20792129}
(\texttt{isSupplementTo} the pre-registration, \texttt{continues} the C1/C2
records); the pre-analysis SHA-256 integrity manifest is committed as
\texttt{c1-c2-hotel-manifest.sha256}. Measuring this workload required general
tooling fixes (a wget-capable readiness-probe pod, a static-topology fallback for
the cross-node fraction on Consul/gRPC services, and sustained warm-up load
through the readiness gate), none of which alter the experimental design or the
system under test.
